---
title: "R Data Science - Mising Data"
output: html_notebook
authors: Sean Cleveland and Mahdi Belcaid
---
### Working with Missing Data

* Given the pervasiveness of missing values in real data, R provides easy to use functionality for handling missing values.

* When working with missing data, the objectives can be boiled down to:
  + Identifying missing value
  + Filling missing values
  + Filtering rows or column with missing values
  
```{r}
spending_df <- read.csv(file="data/spending_correc_data_t.csv", header=TRUE, sep=",")
row.names(spending_df) <- spending_df$unique_id # set the data frame index to the unique_id column
spending_df$unique_id <- NULL # remove the unique_id column since we have the values in the data frame index
spending_df
```
##### 1. Identifying Missing Values

- This is typically achieved with the  `is.na()` method.
- This method returns `True` if a cell contains a `NaN` value, and returns `False` otherwise.

- When applied to the spending_df, the spending value for `unique_id` CC128705 evaluates `isnull()` to `True`.

```{r}
is.na(spending_df) # returns TRUE if data is missing
```

##### 2. Filtering Out Missing Values - 1

- Various approaches can be used to filter out missing value.


- For instance, you can discard missing values using subsetting.


- For example, you can filter out missing values in the `spending` `Series` using na.omit

```r
spending_df_not_na <- na.omit(spending_df)
spending_df_not_na
```
##### Filling In Missing Values 

- There  are two conventional approaches for filling missing value:
  - Filling the value with a constant 
  - Filling the value dynamically with something computed on the fly.

```r
spending_df_constant <- spending_df
spending_df_constant[is.na(spending_df_constant)] <- 0 #For instance
spending_df_constant
```
##### Filling In Missing Values with a function - like a mean
Often we want to use a better value than a constant so we can replace the NA value with the result of function such as the mean of a column.

- The below ifelse() function returns the  mean value if its first argument is NA. Otherwise, it returns the first argument.  Using this we can replace NA values in speding_df_dynamic with the "spending" column mean (this mean excludes NA values via the na.rm=TRUE)
```r
spending_df_dynamic <- spending_df
spending_df_dynamic$spending <- ifelse(is.na(spending_df_dynamic$spending), mean(spending_df_dynamic$spending, na.rm=TRUE), spending_df_dynamic$spending)
spending_df_dynamic
```

```r
temp_spending_df <- spending_df

```


